---
title: "data_import"
author: "Amadeia Rector"
date: "9/20/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Import FAS csv files to use as examples.
```{r}
litters_data = read_csv(file = "./wrangle_data/FAS_litters.csv")
litters_data = janitor::clean_names(litters_data)
```

## Import FAS_pups csv files
```{r}
FAS_pups_data = read_csv(file = "./wrangle_data/FAS_pups.csv", col_types = "ciiiii")
FAS_pups_data = janitor::clean_names(FAS_pups_data)
skimr::skim(FAS_pups_data)
```

## Select variables
-selects columns
```{r}
select(litters_data, group, litter, gd0_weight, pups_survive)
```
... by specifying range:

```{r}
select(litters_data, gd_of_birth:pups_survive)
```

```{r}
select(litters_data, -litter_number)
```

#selecting variables but renaming
```{r}
select(litters_data, group, litter_number, gest_day_0_weight = gd0_weight)
rename(litters_data,gest_day_0_weight=gd0_weight)
```

## this helps put litter_number ahead
```{r}
select(litters_data, starts_with("pup"))
select(litters_data, litter_number, everything())
```

## Filter observations
Like to eliminate some rows from data set -- can you identify a feature from some variable

Do some filtering...

```{r}
filter(litters_data, gd0_weight < 25)
filter(litters_data, gd0_weight >= 25)
filter(litters_data, pups_born_alive == 8)

filter(litters_data, !is.na(gd0_weight))
filter(litters_data, group == "Low7" | group == "Low8")
filter(litters_data, group %in% c("Low7", "Low8"))
filter(litters_data, pups_born_alive %in% 4:7)
```

```{r}
filter(FAS_pups_data, sex == 1)
```

## Mutate

Create variables based on functions of another variable. overriding existing group variable to a new version 
of group

```{r}
mutate(litters_data,
       wt_gain = gd18_weight - gd0_weight,
       group = tolower(group)
)
```

```{r}
mutate(litters_data,
       wt_gain = gd18_weight - gd0_weight,
       wt_gain_squared = wt_gain^2)
```

## Arrange- not changing anything just putting in new order
```{r}
arrange(litters_data,gd0_weight)
arrange(litters_data,pups_born_alive,gd0_weight)
```

## Piping


